Figures and results supporting main manuscript

### 2020-12-17 analysis/reports/figure_s3.Rmd
reorganizing reports and remove selfsim metrics fromcodecov as they are supplementary

---
title: "Figures and results supporting main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

>>>>

---
title: ""
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

### 2020-12-17 analysis/reports/figure_s5.Rmd
reorganizing reports and remove selfsim metrics fromcodecov as they are supplementary

---
title: "Figures and results supporting main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

@@

## Figure S3. Final dataset in S and N space

```{r final dataset s and n space}

s3_fig1 <- ggplot(all_di, aes(x = log_s0, y = log_n0, color = Dataset)) +
  geom_point(alpha = .1) +
  geom_point(data = filter(all_di, log_nparts > log(50)), alpha = .4) +
  theme_bw() +
  scale_color_viridis_d(end = .9) +
  ggtitle("Communities by S and N") +
  xlab("Log species richness (S)") +
  ylab("Log total abundance (N)")

plot(s3_fig1)
# ggplot(all_di, aes(x = log_s0, y = log_n0, color = Dataset)) +
#   geom_point(alpha = .05) +
#   geom_point(data = filter(all_di, log_nparts > log(50)), alpha = .4) +
#   theme_bw() +
#   scale_color_viridis_d(end = .9) +
#   ggtitle("Communities by S and N") +
#   facet_wrap(vars(Dataset), scales = "free") +
#   theme(legend.position = "bottom")
# # geom_vline(xintercept = log(40)) +
# # geom_hline(yintercept = log(1000))

pdf("figure_s3.pdf")
plot(s3_fig1)
dev.off()
```

# Table S4. Proportion of sites with highly skewed or uneven SADs

```{r results table}

overall_skew_results <- all_di %>%
  filter(s0 > 2, skew_unique > 20) %>%
  mutate(high_skew = skew_percentile_excl > 95) %>%
  group_by(Dataset) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results <- all_di %>%
  filter(simpson_unique > 20) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(Dataset) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

overall_results <- left_join(overall_skew_results, overall_simpson_results) %>%
  rename("Proportion of communities with skewness above 95th percentile" = proportion_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = proportion_even_low,
         "Number of communities analyzed for evenness" = nsites_even)

overall_results

overall_skew_results2 <- all_di %>%
  mutate(Dataset = ifelse(dat == "fia", "Forest Inventory and Analysis", "Other datasets")) %>%
  filter(s0 > 2, skew_unique > 20) %>%
  mutate(high_skew = skew_percentile_excl > 95) %>%
  group_by(Dataset) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results2 <- all_di %>%
  filter(simpson_unique > 20) %>%
  mutate(Dataset = ifelse(dat == "fia", "Forest Inventory and Analysis", "Other datasets")) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(Dataset) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

overall_results2 <- left_join(overall_skew_results2, overall_simpson_results2) %>%
  rename("Proportion of communities with skewness above 95th percentile" = proportion_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = proportion_even_low,
         "Number of communities analyzed for evenness" = nsites_even)

#overall_results2

@@

# Table S6. FIA comparison results

```{r}

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia = filter(all_di, dat == "fia")

small_di <- all_di %>%
  filter(s0 <= fia_max_s,
         n0 <= fia_max_n) %>%
  mutate(fia_yn = ifelse(dat == "fia", "fia", "other datasets"),
         right_corner = FALSE)

for(i in 1:nrow(small_di)) {
  if(small_di$dat[i] != "fia") {
    fia_match_s0 = filter(fia, s0 >= small_di$s0[i])

    max_n0_for_this_s0 = max(fia_match_s0$n0)

    small_di$right_corner[i] = small_di$n0[i] > max_n0_for_this_s0

  }
}

small_di <- small_di %>%
  filter(!right_corner)
not_fia <- filter(small_di, fia_yn != "fia")
fia = filter(small_di, dat == "fia")

small_di_s_n <- small_di %>%
  select(s0, n0) %>%
  distinct() %>%
  mutate(in_fia = FALSE,
         in_not_fia = FALSE)

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)
  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  small_di_s_n$in_fia[i] = nrow(fia_matches) > 0 
  small_di_s_n$in_not_fia[i] = nrow(not_fia_matches) > 0

}

small_di_s_n <- filter(small_di_s_n, in_fia, in_not_fia)

set.seed(1977)

subsamples <- list()

for(i in 1:nrow(small_di_s_n)) {

  this_s0 = small_di_s_n$s0[i]
  this_n0 = small_di_s_n$n0[i]

  fia_matches = filter(fia, s0 == this_s0, n0 == this_n0)

  not_fia_matches = filter(not_fia, s0 == this_s0, n0 == this_n0)

  n_to_draw = min(nrow(fia_matches), nrow(not_fia_matches))

  fia_draw = sample.int(n = nrow(fia_matches), size = n_to_draw, replace = F)
  not_fia_draw = sample.int(n = nrow(not_fia_matches), size = n_to_draw, replace = F)

  subsamples[[i]] <- (bind_rows(fia_matches[fia_draw,], not_fia_matches[not_fia_draw, ]))

}

sub_di <- bind_rows(subsamples) %>%
  mutate(Dataset = ifelse(fia_yn == "fia", "FIA", "Other datasets"))

sub_di %>%
  filter(s0 > 2) %>%
  group_by_all %>%
  mutate(skew_percentile_excl = ifelse(skew_unique >= 20, skew_percentile_excl, NA),
         simpson_percentile = ifelse(simpson_unique >= 20, simpson_percentile, NA)) %>%
  ungroup() %>%
  group_by(Dataset) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95, na.rm = T),
            prop_even_low = mean(simpson_percentile < 5, na.rm = T),
            nsites_skew = sum(!is.na(skew_percentile_excl)),
            nsites_even = sum(!is.na(simpson_percentile))) %>%
   rename("Proportion of communities with skewness above 95th percentile" = prop_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = prop_even_low,
         "Number of communities analyzed for evenness" = nsites_even)

>>>>

---
title: ""
output: 
    github_document:
       df_print: kable
---

@@

@@

### 2020-12-17 analysis/reports/table_s6.Rmd
reorganizing reports and remove selfsim metrics fromcodecov as they are supplementary

---
title: "Figures and results supporting main manuscript"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

@@

## Figure S3. Final dataset in S and N space

```{r final dataset s and n space}

s3_fig1 <- ggplot(all_di, aes(x = log_s0, y = log_n0, color = Dataset)) +
  geom_point(alpha = .1) +
  geom_point(data = filter(all_di, log_nparts > log(50)), alpha = .4) +
  theme_bw() +
  scale_color_viridis_d(end = .9) +
  ggtitle("Communities by S and N") +
  xlab("Log species richness (S)") +
  ylab("Log total abundance (N)")

plot(s3_fig1)
# ggplot(all_di, aes(x = log_s0, y = log_n0, color = Dataset)) +
#   geom_point(alpha = .05) +
#   geom_point(data = filter(all_di, log_nparts > log(50)), alpha = .4) +
#   theme_bw() +
#   scale_color_viridis_d(end = .9) +
#   ggtitle("Communities by S and N") +
#   facet_wrap(vars(Dataset), scales = "free") +
#   theme(legend.position = "bottom")
# # geom_vline(xintercept = log(40)) +
# # geom_hline(yintercept = log(1000))

pdf("figure_s3.pdf")
plot(s3_fig1)
dev.off()
```

# Table S4. Proportion of sites with highly skewed or uneven SADs

```{r results table}

overall_skew_results <- all_di %>%
  filter(s0 > 2, skew_unique > 20) %>%
  mutate(high_skew = skew_percentile_excl > 95) %>%
  group_by(Dataset) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results <- all_di %>%
  filter(simpson_unique > 20) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(Dataset) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

overall_results <- left_join(overall_skew_results, overall_simpson_results) %>%
  rename("Proportion of communities with skewness above 95th percentile" = proportion_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = proportion_even_low,
         "Number of communities analyzed for evenness" = nsites_even)

overall_results

overall_skew_results2 <- all_di %>%
  mutate(Dataset = ifelse(dat == "fia", "Forest Inventory and Analysis", "Other datasets")) %>%
  filter(s0 > 2, skew_unique > 20) %>%
  mutate(high_skew = skew_percentile_excl > 95) %>%
  group_by(Dataset) %>%
  summarize(proportion_skew_high = mean(high_skew),
            nsites_skew = dplyr::n())

overall_simpson_results2 <- all_di %>%
  filter(simpson_unique > 20) %>%
  mutate(Dataset = ifelse(dat == "fia", "Forest Inventory and Analysis", "Other datasets")) %>%
  mutate(low_even = simpson_percentile < 5) %>%
  group_by(Dataset) %>%
  summarize(proportion_even_low = mean(low_even),
            nsites_even = dplyr::n())

overall_results2 <- left_join(overall_skew_results2, overall_simpson_results2) %>%
  rename("Proportion of communities with skewness above 95th percentile" = proportion_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = proportion_even_low,
         "Number of communities analyzed for evenness" = nsites_even)

#overall_results2

# Figure S5. Breadth index by size of FS

```{r 95 intervals}
s3_fig2 <- gridExtra::grid.arrange(grobs = list(
  ggplot(filter(all_di, s0 > 2, skew_unique > 20), aes(x = log_nparts, y= skew_95_ratio_1t)) +
  geom_point(alpha = .1) +
  theme_bw() +
  xlim(0, 100) +
  ylim(0,1)+
  xlab("Log number of elements in the feasible set") +
  ylab("Breadth index") +
  ggtitle("Breadth index (skewness)"),

ggplot(filter(all_di, simpson_unique > 20), aes(x = log_nparts, y= simpson_95_ratio_1t)) +
  geom_point(alpha = .1) +
  theme_bw() +
  xlim(0, 100) +
  ylim(0, 1)+
  xlab("Log number of elements in the feasible set") +
  ggtitle("Breadth index (evenness)") +
  ylab("Breadth index")
), ncol = 2)

pdf("figure_s5.pdf")
plot(s3_fig2)
dev.off()
```

# Table S6. FIA comparison results

```{r}

@@

sub_di %>%
  filter(s0 > 2) %>%
  group_by_all %>%
  mutate(skew_percentile_excl = ifelse(skew_unique >= 20, skew_percentile_excl, NA),
         simpson_percentile = ifelse(simpson_unique >= 20, simpson_percentile, NA)) %>%
  ungroup() %>%
  group_by(Dataset) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95, na.rm = T),
            prop_even_low = mean(simpson_percentile < 5, na.rm = T),
            nsites_skew = sum(!is.na(skew_percentile_excl)),
            nsites_even = sum(!is.na(simpson_percentile))) %>%
   rename("Proportion of communities with skewness above 95th percentile" = prop_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = prop_even_low,
         "Number of communities analyzed for evenness" = nsites_even)

>>>>

---
title: ""
output: 
    word_document:
       df_print: kable
---

@@

# Table S6. Proportion of sites with highly skewed or uneven SADs for a subset of FIA communities compared to matching communities from other datasets

```{r, include = F, echo = F}

@@

```
```{r}
sub_di %>%
  filter(s0 > 2) %>%
  group_by_all %>%
  mutate(skew_percentile_excl = ifelse(skew_unique >= 20, skew_percentile_excl, NA),
         simpson_percentile = ifelse(simpson_unique >= 20, simpson_percentile, NA)) %>%
  ungroup() %>%
  group_by(Dataset) %>%
  summarize(prop_skew_high = mean(skew_percentile_excl > 95, na.rm = T),
            prop_even_low = mean(simpson_percentile < 5, na.rm = T),
            nsites_skew = sum(!is.na(skew_percentile_excl)),
            nsites_even = sum(!is.na(simpson_percentile))) %>%
   rename("Proportion of communities with skewness above 95th percentile" = prop_skew_high,
         "Number of communities analyzed for skewness" = nsites_skew,
         "Proportion of communities with evenness below 5th percentile" = prop_even_low,
         "Number of communities analyzed for evenness" = nsites_even)
